WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2015-07-07 ms.tex
Add caption

This paper is organized as follows. In Section~\ref{sec:detection}, we describe the \ac{GW} signal model and sketch the standard detection algorithm, the matched filter bank. In Section~\ref{sec:parameter-estimation}, we describe Bayesian inference formalism and the prevailing method for inferring the parameters of detected candidates, \ac{MCMC} sampling. In Section~\ref{sec:bayestar-likelihood}, we propose the \ac{BAYESTAR} likelihood as a model for the uncertainty in the matched\nobreakdashes-filter parameter estimates, and discuss its relationship to and consistency with the likelihood for the full \ac{GW} data. In Section~\ref{sec:prior}, we describe the input to \ac{BAYESTAR} supplied by the detection pipeline, and the prior distribution on parameters. In Section~\ref{sec:marginal-posterior}, we explain the integration scheme by which the posterior probability distribution is calculated for a given sky location. In Section~\ref{sec:adaptive-sampling}, we show a scheme whereby the sky posterior is sampled on an adaptive HEALPix grid. In Section~\ref{sec:run-time}, we report the running time of the algorithm on the hardware available on the \ac{LIGO} Data Grid. Finally, in Section~\ref{sec:case-study}, we quantify the sky localization performance on a comprehensive set of simulated \ac{GW} events.

@@

Our sampling of the sky relies completely on the \acl{HEALPix} (\acs{HEALPix}), a special data structure designed for all\nobreakdashes-sky maps. \ac{HEALPix} divides the sky into equal\nobreakdashes-area pixels. There is a hierarchy of \ac{HEALPix} resolutions. A \ac{HEALPix} resolution may be designated by its order $N$. The $N=0$th order or base tiling has 12 pixels. At every successive order, each tile is subdivided into four new tiles. A resolution may also be referred to by the number of subdivisions along each side of the base tiles, $N_\mathrm{side} = 2^N$. There are $N_\mathrm{pix} = 12 {N_\mathrm{side}}^2$ pixels at any given resolution. The \ac{HEALPix} projection uniquely specifies the coordinates of the center of each pixel by providing a mapping from the resolution and pixel index $(N_\mathrm{side}, i_\mathrm{pix})$ to right ascension and declination $(\alpha, \delta)$.

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{adaptive_mesh}
    \caption{\label{fig:adaptive-mesh}}
\end{figure}

>>>>

This paper is organized as follows. In Section~\ref{sec:detection}, we describe the \ac{GW} signal model and sketch the standard detection algorithm, the matched filter bank. In Section~\ref{sec:parameter-estimation}, we describe Bayesian inference formalism and the prevailing method for inferring the parameters of detected candidates, \ac{MCMC} sampling. In Section~\ref{sec:bayestar-likelihood}, we propose the \ac{BAYESTAR} likelihood as a model for the uncertainty in the matched\nobreakdashes-filter parameter estimates, and discuss its relationship to and consistency with the likelihood for the full \ac{GW} data. In Section~\ref{sec:prior}, we describe the input to \ac{BAYESTAR} supplied by the detection pipeline, and the prior distribution on parameters. In Section~\ref{sec:marginal-posterior}, we explain the integration scheme by which the posterior probability distribution is calculated for a given sky location. In Section~\ref{sec:adaptive-sampling}, we show a scheme whereby the sky posterior is sampled on an adaptive \acs{HEALPix} grid. In Section~\ref{sec:run-time}, we report the running time of the algorithm on the hardware available on the \ac{LIGO} Data Grid. Finally, in Section~\ref{sec:case-study}, we quantify the sky localization performance on a comprehensive set of simulated \ac{GW} events.

@@

Our sampling of the sky relies completely on the \acl{HEALPix} (\acsu{HEALPix}), a special data structure designed for all\nobreakdashes-sky maps. \ac{HEALPix} divides the sky into equal\nobreakdashes-area pixels. There is a hierarchy of \ac{HEALPix} resolutions. A \ac{HEALPix} resolution may be designated by its order $N$. The $N=0$th order or base tiling has 12 pixels. At every successive order, each tile is subdivided into four new tiles. A resolution may also be referred to by the number of subdivisions along each side of the base tiles, $N_\mathrm{side} = 2^N$. There are $N_\mathrm{pix} = 12 {N_\mathrm{side}}^2$ pixels at any given resolution. The \ac{HEALPix} projection uniquely specifies the coordinates of the center of each pixel by providing a mapping from the resolution and pixel index $(N_\mathrm{side}, i_\mathrm{pix})$ to right ascension and declination $(\alpha, \delta)$.

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{adaptive_mesh}
    \caption{\label{fig:adaptive-mesh}An example multi\nobreakdashes-resultion \ac{HEALPix} mesh arising from the \ac{BAYESTAR} sampling scheme. This is event 18951 from \cite{FirstTwoYears}.}
\end{figure}
